\begin{abstract}

This is an abstract.
It is currently abstract.
It is way too abstract for you to understand as it is now.

\paragraph
As of such, for your enjoyable reading, we will provide with the lyrics from the Lion King song "Be Prepared":

\paragraph
Scar:
I know that your powers of retention
Are as wet as a warthog's backside
But thick as you are, pay attention
My words are a matter of pride
It's clear from your vacant expressions
The lights are not all on upstairs
But we're talking kings and successions
Even you can't be caught unawares
So prepare for a chance of a lifetime
Be prepared for sensational news
A shining new era
Is tiptoeing nearer

Shenzi:
And where do we feature?

Scar:
Just listen to teacher
I know it sounds sordid
But you'll be rewarded
When at last I am given my dues
And injustice deliciously squared
Be prepared!

(talking)
Banzai: Yeah, Be prepared. 
Yeah-heh... we'll be prepared, heh.
...For what?

Scar: For the death of the king.

Banzai: Why? Is he sick?

Scar: No, fool, we're going to kill him. And Simba too.

Shenzi: Great idea! Who needs a king?

Shenzi (and then Banzai):
No king! No king! la--la-la--la-laa-laa!

Scar: Idiots! There will be a king!

Banzai: Hey, but you said, uh...

Scar: I will be king! ...Stick with me, and
you'll never go hungry again!

Shenzi and Banzai: Yaay! All right! Long live the king!

All Hyenas: Long live the king! Long live the king!

(Full song again)
Hyenas: (In tight, crisp phrasing and diction
It's great that we'll soon be connected.
With a king who'll be all-time adored.

Scar: Of course, quid pro quo, you're expected
To take certain duties on board
The future is littered with prizes
And though I'm the main addressee
The point that I must emphasize is
You won't get a sniff without me!
So prepare for the coup of the century
(Oooh!)
Be prepared for the murkiest scam
(Oooh... La! La! La!)
Meticulous planning
(We'll have food!)
Tenacity spanning
(Lots of food)
Decades of denial
(We repeat)
Is simply why I'll
(Endless meat)
Be king undisputed
(Aaaaaaah...)
Respected, saluted
(...aaaaaaah...)
And seen for the wonder I am
(...aaaaaaah!)
Yes, my teeth and ambitions are bared
(Oo-oo-oo-oo-oo-oo-oo)
Be prepared!

All:
Yes, our teeth and ambitions are bared
Be prepared!


\end{abstract}

