\chapter{The Assignment}

\section{Original Assignment Text}

This project aims to develop a bitcoin mining accelerator that will ultimately be
used in the single-ISA, many-core, heterogeneous computing platform SHMAC. Bitcoin
mining is, at its core, a SHA256 hashing problem, so part of the assignment will
be to keep the interface generic enough such that other cryptographic algorithms
can be readily developed.

This part of the project will focus on building an end-to-end and fully functional
hardware and software system that is able to fully participate as a miner in the
network (that is, mine Bitcoins and communicate with peers).

The energy efficiency of the hardware implementation should be evaluated against that
of a general-purpose CPU and should constitute a major aspect of the report.

\section{Comments on the Assignment Text}

Because of recent developments within the field of bitcoin mining, the assignment
is basically obsolete. Modern bitcoin hardware\todo{Add section reference} is typically
systems with custom-designed ASIC chips for the best performance and power efficiency.

The assignment then boils down to: is bitcoin mining possible using SHMAC? Does the
mining process benefit from heterogenity and what performance gains are possible
using a dedicated SHA256 accelerator and a DMA?

Other issues relevant are how the system scales; given enough processing cores,
could SHMAC compare to dedicated bitcoin mining systems in power efficiency or
computing power?

\todo{Add lots of references to other parts in the text.}

